---
title: "Lab 5 Assignment"
author: "Beau Ayers"
date: "2021-02-23"
output: github_document
---
```{r}
# Load Libraries ----------------------------------------------------------

library(tidyverse)

# Print Data --------------------------------------------------------------

iris
iris <- as_tibble(iris)
iris
```

Rename each variable so that it is all lower-case and uses an underscore _ instead of a period . in the name (the recommended coding style in the tidyverse style guide). Print the resulting table.

```{r}
iris_r <- rename(iris, sepal_length = Sepal.Length,
       sepal_width = Sepal.Width,
       petal_length = Petal.Length,
       petal_width = Petal.Width,
       species = Species
       )
iris_r
```

Convert the four numerical variables from cm to mm by multiplying by 10. Print the resulting table.

```{r}
iris_cm <- mutate(iris_r, sepal_length = sepal_length * 10,
                  sepal_width = sepal_width * 10,
                  petal_length = petal_length * 10,
                  petal_width = petal_width * 10)
iris_cm
```

Calculate sepal area and petal area (area is equal to length multiplied by width). Print a table with only the variables sepal area, petal area, and species.

```{r}
iris_area <- mutate(iris_cm, sepal_area = sepal_length * sepal_width,
                    petal_area = petal_length * petal_width) %>% 
  select(petal_area, sepal_area, species)

iris_area
```

Calculate the following statistics for the entire dataset from the sepal length variable and print the resulting table:sample size, maximum value, minimum value,
range, median, first quartile (q1), third quartile (q2), 
inter-quartile range (iqr)

```{r}
summarize(
  iris_cm,
  sampl_size = n(),
  max = max(sepal_length),
  min = min(sepal_length),
  range = max - min,
  med = median(sepal_length),
  q1 = quantile(sepal_length, probs = 0.25),
  q3 = quantile(sepal_length, probs = 0.75),
  iqr = IQR(sepal_length)
)

```

Calculate the following statistics for each species from the petal width variable and print the resulting table: sample size, mean, standard deviation, variance,
standard error of the mean, approximate 95% confidence interval

```{r}
iris_sum <-
  iris_cm %>%
  group_by(species) %>%
  summarize(
    sampl_size = n(),
    mean_w = mean(petal_width),
    str_dev = sd(petal_width),
    var = var(petal_width),
    sem = mean(petal_width) / sqrt(n()),
    ci_upper = mean_w + 2 * sem,
    ci_lower = mean_w - 2 * sem,
  )

iris_sum
```

Visualize the relationship between petal width and species using a strip plot.

```{r}
ggplot(data = iris_cm) +
  geom_jitter(mapping = aes(x = species, y = petal_width))
```

Starting with the previous graph, add the mean and 95% confidence interval for each species

```{r}
ggplot(data = iris_cm) +
  geom_jitter(mapping = aes(x = species, y = petal_width)) +
  geom_crossbar(
    data = iris_sum,
    mapping = aes(x = species, y = mean_w, ymax = ci_upper, ymin = ci_lower),
    color = "darkorchid4"
  )
```

Visualize the relationship between petal length, petal width, and species using a scatterplot. Map the two numerical variables to the x and y axes and map species to the color and shape aesthetics.

```{r}
ggplot(data = iris_cm) +
  geom_point(mapping = aes(x = petal_length,
                          y = petal_width,
                          color = species)) 
```

Here is my session info 

```{r}
sessioninfo::session_info()
```

